---
title: "DSS Manuscript"
output: html_notebook
---

#Libraries
```{r}
library("phyloseq")
library("tidyverse")
library("qiime2R")
library("ggpubr")
library("RColorBrewer")
library("data.table")
library("ggridges")
library(vegan)
library(cowplot)
library(rstatix)
library(e1071)
library(gridExtra)
library(DT)
library("ape")
library("biomformat")
library("dplyr")
library(rstatix)
library("car")
library(ggplot2)
library(ggpubr)
library(rstatix)
library(vegan)
```

#Bring in DSS that was merged in qiime, with all filtering already done. filtered out sequences that were present in sheath fluid but absent in pellet (from the ProperFiltering folder)
```{r}
ps <- qza_to_phyloseq("TableDSSAllPF-filt2.qza","rooted-tree.qza","taxonomy.qza")

# #or if the above doesn't work:
# tmp <- tempdir()
# features<-read_qza('TableDSSAllPF-filt2.qza', tmp=tmp)$data
# taxonomy<-read_qza('taxonomy.qza', tmp=tmp)$data
# taxonomy<-parse_taxonomy(taxonomy)
# taxonomy<-as.matrix(taxonomy)
# tree<-read_qza('rooted-tree.qza', tmp=tmp)$data


my_otu <- otu_table(features, taxa_are_rows = T)
my_tree <- phy_tree(tree)
my_taxa <- tax_table(taxonomy)
metadata<-read_tsv("metadata-allDSS.tsv") 
my_meta <- sample_data(metadata) 
sample_names(my_meta) <- metadata$sampleid
ps2 <- phyloseq(my_otu, my_tree, my_taxa, my_meta)
tax_table(ps2)
```


#Removing Contaminants
```{r}
ps2 <- subset_samples(ps2, Physiology != "H-")
SF <- subset_samples(ps2, Physiology == "SF")
SF <- prune_taxa(taxa_sums(SF) > 0, SF)
SFnames <- taxa_names(SF) #taxa that are present in sheath fluid, 271 taxa present in the sheath fluid
DNAps2 <- subset_samples(ps2, Physiology == "PELLET") #867 taxa in unsorted, the original pellet
DNAps2 <- prune_taxa(taxa_sums(DNAps2) < 1, DNAps2) #everything that is absent in the original pellet (taxa sum is less than 0) 187
DNAps2names <- taxa_names(DNAps2) #everything absent in the pellet, which is 187 taxa!
#Now I want to find the intersection between SFnames and DNAps2names (ie found in SF, but absent from DNA, as in the name DNAps2names)
Taxa_to_filter <- intersect(SFnames, DNAps2names)  # 7 taxa
#there are 7 taxa that are present in sheath fluid but absent from DNA  (remember i did filtering in  qiime)
Taxa_to_keep <- setdiff(taxa_names(ps2), Taxa_to_filter)
length(Taxa_to_keep)  
ps2_NoSF <- prune_taxa(Taxa_to_keep, ps2) #860 taxa, getting rid of the 7

ps0 <- subset_taxa(ps2_NoSF, !is.na(Phylum) & !Phylum %in% c("", "uncharacterized")) #Now at 11772 taxa, removed 37
ps4 <- filter_taxa(ps0, function(x) sum(x > 4) > (0.02*length(x)), TRUE) #512 taxa (from 11772! that's only 4% kept)
Phys = c("HNA", "LNA", "PI", "H+", "H-", "PELLET")
DSS <- subset_samples(ps4, Physiology %in% Phys) #479 taxa, 777 samples
colnames(sample_data(DSS))
DSS3 <- subset_samples(DSS, Experiment == "DSS3") #283 samples
DSS4 <- subset_samples(DSS, Experiment == "4") #245 samples
DSS2 <- subset_samples(DSS, Experiment == "DSS2-1" | Experiment == "DSS2-2"| Experiment == "DSS2") #249 sample
```

#Cleanup 
```{r}
DSS <- subset_samples(DSS, !is.na(Physiology))
DSS <- subset_samples(DSS, !is.na(Day))
DSSa <- subset_samples(DSS, Mouse != "nH")
#DSSx <- subset_samples(DSSa, Day != "1"& Day != "2"& Day != "8"& Day != "9")

DSSx <- subset_samples(DSSa, Controls != "control" | is.na(Controls) ) 
#Glom to species level since multiple DADA2 runs are combined together
#DSSx <- tip_glom(DSSx, h = 0.1)
#glomming the tips to 0.1 gets it down to 108 taxa from 479, retainign 22%, genus level is 88 taxa,

#Turn BS+  to BCpos 
new_sample_data <- as(sample_data(DSSx), "data.frame") %>%
  mutate(Physiology = ifelse(Physiology == "H+", "BCpos", Physiology)) %>%
  mutate(Physiology = ifelse(Physiology == "H++", "BCpospos", Physiology))
new_sample_data <- sample_data(new_sample_data)
sample_names(new_sample_data) <- sample_names(sample_data(DSSx))
sample_data(DSSx) <- new_sample_data
```

#Figure 5 Barplots
```{r}
sample_data(DSSx)$Day = as.numeric(sample_data(DSSx)$Day)
DSSPhyl <- tax_glom(DSSx, "Phylum")
DSSPhylRA <- transform_sample_counts(DSSPhyl, function(x) x / sum(x) )
DSSRAPhylumNC  <- subset_samples(DSSPhylRA , !Physiology %in% Controls) 

df <- DSSRAPhylumNC %>% psmelt() 

df$Physiology <- factor(df$Physiology, levels = c("PELLET", "BCpos", "HNA", "LNA", "PI")  )

aggregate(Abundance ~ Day + Physiology + Phylum, data = df, mean) %>%
  ggplot(aes(y = Abundance, x = Day)) +
  geom_bar(aes(fill = Phylum), stat='identity') + ylab("Relative Abundance") + scale_y_continuous(name = "Relative Abundance", breaks = c(0,0.5,1)) +
  facet_grid(rows = vars(Physiology)) +theme(text = element_text(size = 14)) + theme_bw() + scale_fill_brewer(palette = "Accent")
ggsave("PhylaLevelTaxaBarplot.pdf", width = 6, height = 7)


#Genus
DSSRA <- transform(DSSx, "compositional")
DSSRAGenus <- tax_glom(DSSRA, "Genus")

Top10Genus <- names(sort(taxa_sums(DSSRAGenus), TRUE)[1:10])
Top10GenusDSS <- prune_species(Top10Genus, DSSRAGenus)
#Take what is not in top 10 and name as other
AllGenus <- taxa_names(DSSRAGenus)
OtherGenera <- setdiff(AllGenus, Top10Genus) #"c8d2b3167ef685d221379c95b2553a66"
DSS_TopGenus <- merge_taxa(DSSRAGenus, OtherGenera, 1)
DSSRAGenusNC  <- subset_samples(DSS_TopGenus , !Physiology %in% Controls) 
#remove physiology is na
DSSRAGenusNC <- subset_samples(DSSRAGenusNC, !is.na(Physiology))
#remove mous is nH
DSSRAGenusNC <- subset_samples(DSSRAGenusNC, Mouse != "nH")

df <- DSSRAGenusNC %>% psmelt()

df$Physiology <- factor(df$Physiology, levels = c("PELLET", "BCpos", "HNA", "LNA", "PI") )

unique(df$Genus)

aggregate(Abundance ~ Day + Physiology + Genus, data = df, mean) %>%
  ggplot(aes(y = Abundance, x = Day)) +
  geom_bar(aes(fill = Genus), stat='identity') + ylab("Relative Abundance") + 
  scale_y_continuous(name = "Relative Abundance", breaks = c(0,0.5,1)) +
  facet_grid(rows = vars(Physiology)) +theme(text = element_text(size = 14)) + theme_bw()
ggsave("GenusLevelTaxaBarplot.pdf", width = 7.2, height = 7)
```

#Permanova functions
```{r}
#By Day
Permanova_WU <- function(subset_phys_ps) {
  counts = otu_table(subset_phys_ps)
  tree = phy_tree(subset_phys_ps)
  rbiom_weighted.rlog = rbiom::unifrac(counts, weighted=TRUE, tree=tree)
  adonis(rbiom_weighted.rlog ~ Day,
       data = as(sample_data(subset_phys_ps), "data.frame"))
}

#By DiseaseState
Permanova_WU_DS <- function(subset_phys_ps) {
  counts = otu_table(subset_phys_ps)
  tree = phy_tree(subset_phys_ps)
  rbiom_weighted.rlog = rbiom::unifrac(counts, weighted=TRUE, tree=tree)
  adonis(rbiom_weighted.rlog ~ DiseaseState,
       data = as(sample_data(subset_phys_ps), "data.frame"))
}

#By Physiology
Permanova_WU_Phys <- function(subset_phys_ps) {
  counts = otu_table(subset_phys_ps)
  tree = phy_tree(subset_phys_ps)
  rbiom_weighted.rlog = rbiom::unifrac(counts, weighted=TRUE, tree=tree)
  adonis(rbiom_weighted.rlog ~ Physiology,
       data = as(sample_data(subset_phys_ps), "data.frame"))
}
```

#Permanova of baseline
```{r}
DSSx.rel <- microbiome::transform(DSSx, "compositional")

#get just baseline
Baseline <- subset_samples(DSSx.rel, DiseaseState == "baseline")
#remove rows that have NA in mouse
Baseline <- subset_samples(Baseline, !is.na(Mouse))

#ASV
counts = otu_table(Baseline)
tree = phy_tree(Baseline)
metadata <- as(sample_data(Baseline), "data.frame")
adonis(rbiom::unifrac(counts, weighted=TRUE, tree=tree) ~ Physiology*Day, data = metadata)

#Phylum
Baseline.p <- tax_glom(Baseline, "Phylum")
counts = otu_table(Baseline.p)
tree = phy_tree(Baseline.p)
metadata <- as(sample_data(Baseline.p), "data.frame")
adonis(rbiom::unifrac(counts, weighted=TRUE, tree=tree) ~ Physiology*Day, data = metadata)

#Genus
Baseline.g <- tax_glom(Baseline, "Genus")
counts = otu_table(Baseline.g)
tree = phy_tree(Baseline.g)
metadata <- as(sample_data(Baseline.g), "data.frame")
adonis(rbiom::unifrac(counts, weighted=TRUE, tree=tree) ~ Physiology*Day, data = metadata)
```

#Table 1 Permanova by disease state
```{r}
# are physiological groups distinct from one another (**add p value of permanova at each disease state

#subset by disease state
Baseline <- subset_samples(DSSx.rel, DiseaseState == "baseline")
Presympt <- subset_samples(DSSx.rel, DiseaseState == "pre-symptoms")
Symp <- subset_samples(DSSx.rel, DiseaseState == "symptoms")
Recovery <-  subset_samples(DSSx.rel, DiseaseState == "recovery")
W1 <- subset_samples(DSSx.rel, DiseaseState == "W1")
W2 <- subset_samples(DSSx.rel, DiseaseState == "W2")

Permanova_WU_Phys(Baseline)

Permanova_WU_Phys(Presympt)

Permanova_WU_Phys(Symp)

Permanova_WU_Phys(Recovery)

Permanova_WU_Phys(W1)

Permanova_WU_Phys(W2)

#Disease state as a function of physiology 
counts = otu_table(DSSx.rel)
tree = phy_tree(DSSx.rel)
metadata <- as(sample_data(DSSx.rel), "data.frame")
adonis(rbiom::unifrac(counts, weighted=TRUE, tree=tree) ~ Physiology*DiseaseState, data = metadata)
```

#Figure 3 WUF
```{r}
HNA <- subset_samples(DSSx.rel, Physiology == "HNA")
LNA <- subset_samples(DSSx.rel, Physiology == "LNA")
BCpos <- subset_samples(DSSx.rel, Physiology == "BCpos")
PI <- subset_samples(DSSx.rel, Physiology == "PI")
pellet <- subset_samples(DSSx.rel, Physiology == "PELLET")

PCoADayWU <- function(subset_phys_ps, phys) {
  counts = otu_table(subset_phys_ps)
  tree = phy_tree(subset_phys_ps)
  rbiom_weighted.rlog = rbiom::unifrac(counts, weighted=TRUE, tree=tree)
  pcoa.rlog = ordinate(subset_phys_ps, method="PCoA", distance=rbiom_weighted.rlog)
  plot_ordination(subset_phys_ps, pcoa.rlog, "samples", color="DiseaseState", title = phys) + 
    geom_point(size=3) + theme_bw() +  theme(text = element_text(size = 12)) + 
    scale_color_manual(values=c("#d1ddf1", "#feefc1", "#c9c9c9", "#d98689", "grey58", "grey45")) 
}

pHNA <- PCoADayWU(HNA, "HNA")
pLNA <- PCoADayWU(LNA, "LNA")
pPI <- PCoADayWU(PI, "PI")
pBCpos <- PCoADayWU(BCpos, "BONCAT+")
pPellet <- PCoADayWU(pellet, "Whole community")

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

legend <- get_legend(pHNA)

plot_grid(legend, pPellet + theme(legend.position = "none"), pBCpos + theme(legend.position = "none"), pHNA + theme(legend.position = "none"), pLNA + theme(legend.position = "none"), pPI + theme(legend.position = "none"), nrow = 2)
ggsave("WUFPCoRrelAbunNoTipGlom.pdf" , height = 6.8, width = 9)

Permanova_WU_DS(HNA)
Permanova_WU_DS(LNA)
Permanova_WU_DS(PI)
Permanova_WU_DS(BCpos)
Permanova_WU_DS(pellet)
```


#boxplot function
```{r}
library("reshape2")

plotDistances = function(p, m , s , d , title) {

  # calc distances
  #wu = phyloseq::distance(p, m) #if doing bray curtis
  tree = phy_tree(p)
  counts = otu_table(p)
  wu = rbiom::unifrac(counts, weighted=TRUE, tree=tree)
  wu.m = melt(as.matrix(wu))
  
    # remove self-comparisons
  wu.m = wu.m %>%
    filter(as.character(Var1) != as.character(Var2)) %>%
    mutate(Var1 = as.character(Var1)) %>% mutate(Var2 = as.character(Var2))

  
  # get sample data (S4 error OK and expected)
 sd = sample_data(p)[,c(s, d)]
 
  # combined distances with sample data
  colnames(sd) = c("Var1", "Type1")
  wu.sd = left_join(wu.m, sd, by = "Var1")
  
  
  colnames(sd) = c("Var2", "Type2")
  wu.sd = left_join(wu.sd, sd, by = "Var2")
  
  #only include comparisons within a symptom group
  wu.sd <- filter(wu.sd, Type1 == "baseline")
  #make them in the correct order
  wu.sd$Type2 <- factor(wu.sd$Type2, levels=c("baseline", "pre-symptoms", "symptoms", "recovery"))
 
  # #do the stats
  res.kruskal <- wu.sd %>% kruskal_test(value ~ Type2)
  pwc <-  wu.sd %>% dunn_test(value ~ Type2, p.adjust.method = "BH") %>%  filter(group1 == "baseline")
  pwc <- pwc %>% add_xy_position(x = "Type2")
  print(wu.sd)
  
  # plot
  ggplot(wu.sd, aes(x = Type2, y = value)) +
    theme_bw() +
    geom_boxplot(aes(fill = Type2), outlier.alpha = 0.5) +
    scale_fill_manual(values=c("#d1ddf1", "#feefc1", "#d98689", "#c9c9c9")) +
    theme(axis.text.x=element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    theme(legend.position = "none") +
    ggtitle(title) +
    xlab("") +
    ylab(paste0(m, "")) +
    #stat_compare_means(method = "kruskal.test", label.x = 1.2, label.y = 0.15) +
    stat_pvalue_manual(pwc, hide.ns = TRUE, step.increase  = 0.09, size = 6) 
}
```
#Figure 4a
```{r}
DSSx.rel3 <- subset_samples(DSSx.rel, DiseaseState != "W1")
DSSx.rel3 <- subset_samples(DSSx.rel3, DiseaseState != "W2")

HNA <- subset_samples(DSSx.rel3, Physiology == "HNA" )
LNA <- subset_samples(DSSx.rel3, Physiology == "LNA")
BCpos <- subset_samples(DSSx.rel3, Physiology == "BCpos")
PI <- subset_samples(DSSx.rel3, Physiology == "PI")
Pellet <- subset_samples(DSSx.rel3, Physiology == "PELLET")
p.HNA.BC <- plotDistances(HNA, "Weighted UniFrac", "sampleid", "DiseaseState", "HNA")
p.LNA.BC <- plotDistances(LNA, "Weighted UniFrac", "sampleid", "DiseaseState", "LNA")
p.PI.BC <- plotDistances(PI, "Weighted UniFrac", "sampleid", "DiseaseState", "PI")
p.BCpos.BC <- plotDistances(BCpos, "Weighted UniFrac", "sampleid", "DiseaseState", "BONCAT+")
p.Pellet.BC <- plotDistances(Pellet, "Weighted UniFrac", "sampleid", "DiseaseState", "Whole community")

plot_grid(p.Pellet.BC, p.BCpos.BC,  p.HNA.BC, p.LNA.BC, p.PI.BC, nrow = 2)
ggsave("WUFdistancesboxplotRA-without week12-notipglom.pdf", height = 8, width = 8)
```

#Fig 4b Presymptoms to baseline
```{r}
#plotPairedDistances = function(p, m , s , d , i, phys, e, title) {
  # p = phyloseq object
  # m = distance method
  # s = how you call your sampleids
  # d = disease state column title (how you want to group)
  # i = column title of your indidividauls 
  # phys = column title for physiology 
  # column title to differnetiate between experiments 
  
#wu = phyloseq::distance(p, m) 
DSSx.rel3 <- subset_samples(DSSx.rel3, Physiology != "BCneg")
  tree = phy_tree(DSSx.rel3)
  counts = otu_table(DSSx.rel3)
  wu = rbiom::unifrac(counts, weighted=TRUE, tree=tree)
  wu.m = melt(as.matrix(wu))
  
    # remove self-comparisons
  wu.m = wu.m %>%
    filter(as.character(Var1) != as.character(Var2)) %>%
    mutate(Var1 = as.character(Var1)) %>% mutate(Var2 = as.character(Var2))
  
  # get sample data (S4 error OK and expected)
 sd = sample_data(DSSx.rel3)[,c("sampleid" , "DiseaseState" , "Mouse", "Physiology", "Experiment")]
 
  # combined distances with sample data
  colnames(sd) = c("Var1", "DiseaseState" , "Mouse", "Physiology", "Experiment")
  wu.sd = left_join(wu.m, sd, by = "Var1")
  
  
  colnames(sd) = c("Var2", "DiseaseState" , "Mouse", "Physiology", "Experiment")
  wu.sd = left_join(wu.sd, sd, by = "Var2")
  
  #only include comparisons that are baseline to presymptoms
  wu.sd <- filter(wu.sd, DiseaseState.x == "baseline") # I lose DSS2 because don't have baseline for it
  wu.sd <- filter(wu.sd, DiseaseState.y == "pre-symptoms")
  
  #make them in the correct order
  wu.sd$Physiology.x <- factor(wu.sd$Physiology.x, levels=c("PELLET","BCpos", "HNA", "LNA", "PI" ))

# #do the stats
  res.kruskal <- wu.sd %>% kruskal_test(value ~ Physiology.x)
  pwc <-  wu.sd %>% 
   dunn_test(value ~ Physiology.x, p.adjust.method = "BH") %>% filter(group1 == "PELLET" | group2 == "PELLET")
  pwc <- pwc %>% add_xy_position(x = "Physiology")
   
  # plot
  ggplot(wu.sd, aes(x = Physiology.x, y = value)) +
    theme_bw() +
    geom_boxplot(fill = "#feefc1" ) +
    #scale_color_identity() +
    theme(axis.text.x=element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    ggtitle("Presymptoms to baseline") +
    xlab("") +
    ylab("Weighted UniFrac") +
    #stat_compare_means(method = "kruskal.test", label.x = 1.0, label.y = 1.2) +
    stat_pvalue_manual(pwc, hide.ns = TRUE, size = 6, step.increase = 0.09) +
    theme(text = element_text(size = 14))  


#plotPairedDistances(DSSx.rel3, "bray" , "sampleid" , "DiseaseState" , "Mouse", "Physiology", "Experiment", "BC baseline to presymptoms")
ggsave("baselinepresymptoms BC RA.pdf", width = 4.7, height = 5)
```

#Figure 6a Corncob
```{r}
library(corncob)
library(magrittr)

#Anything that is not weighted unifrac, I should tip glom! Also keep this as raw counts
DSSxtg <- tip_glom(DSSx, h = 0.1) #glomming up high, before was doing 0.1

HNA <- subset_samples(DSSxtg, Physiology == "HNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
LNA <- subset_samples(DSSxtg, Physiology == "LNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
PI <- subset_samples(DSSxtg, Physiology == "PI") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
BCpos <- subset_samples(DSSxtg, Physiology == "BCpos") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
WC <- subset_samples(DSSxtg, Physiology == "PELLET") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()

#controls for the effect of disease state on dispersion
cc <- function(ps, title){
  da <- differentialTest(
  formula = ~ DiseaseState ,
  phi.formula = ~ DiseaseState , 
  formula_null = ~ 1, 
  phi.formula_null = ~ DiseaseState , 
  test = "Wald", boot = FALSE, 
  data = ps,
  fdr_cutoff = 0.05)

which(is.na(da$p)) %>% names
pdf(file=title, width = 10)
plot(da)
}

cc(HNA, "HNA1cc.pdf")
cc(LNA, "LNA1cc.pdf")
cc(PI, "PI1cc.pdf")
cc(BCpos, "BCpos1cc.pdf")
#cc(BCneg, "BCnegcc.pdf")
cc(WC, "WC1cc.pdf")

#sometiems the function breaks and returns an empty pdf, rerun the broken ones here 
daWC <- differentialTest(
  formula = ~ DiseaseState ,
  phi.formula = ~ DiseaseState , 
  formula_null = ~ 1, 
  phi.formula_null = ~ DiseaseState , 
  test = "Wald", boot = FALSE, 
  data = HNA, #change the ps object here
  fdr_cutoff = 0.05)

daWC$significant_taxa

which(is.na(daWC$p)) %>% names

cex <- 2
par(cex.lab=cex, cex.axis=cex, cex.main=cex)
pdf(file="ccHNA.pdf", width = 10, pointsize = 100)
plot(daWC)
```

#Relative Abundance of specific taxa over time, taxa chosen based off peramnova or corncob
```{r}
#Akkermansia
Akkermansia.ps <- subset_taxa(DSSx.rel3, Genus == "Akkermansia") #the only taxa in the phylum verrucomicrobia
Akkermansia <- psmelt(Akkermansia.ps) 
#groupby disease state and physiology
Akkermansia %>% group_by(DiseaseState,Physiology) %>% summarize(mean(Abundance))

#ggplot(Akkermansia, aes(x = Day, y = Abundance)) + geom_point(alpha = 0.5, aes(color = Physiology)) + geom_smooth(method = "loess", aes(color = Physiology)) + theme_bw() + #xlim(c(-3,10)) + ylab("Relative Abundance") + ggtitle("Akkermansia muciniphila")
#ggsave("Amuciiphila.pdf", width = 6.8, height = 5.2)

RA.plot <- function(df, title, pdftitle) {
  df$Physiology <- factor(df$Physiology, levels = c("PELLET", "BCpos", "HNA", "LNA", "PI")  )
  ggplot(df, aes(x = Day, y = Abundance)) + geom_point(alpha = 0.5, aes(color = Physiology)) + geom_smooth(method = "loess", aes(color = Physiology)) +
  theme_bw() + xlim(c(-3,10)) + ylab("Relative Abundance") + ggtitle(title)
ggsave(pdftitle, width = 6.8, height = 5.2)
}

RA.plot(Akkermansia, "Akkermansia", "Akkermansia.pdf")

Eubacterium.ps <- subset_taxa(DSSx.rel3, Genus == "[Eubacterium] coprostanoligenes group")
Eubacterium <- psmelt(Eubacterium.ps)
RA.plot(Eubacterium, "Eubacterium", "Eubacterium.pdf")

Bifidobacterium.ps <- subset_taxa(DSSx.rel3, Genus == "Bifidobacterium")
Bifido <- psmelt(Bifidobacterium.ps)
RA.plot(Bifido, "Bifidobacterium", "Bifidobacterium.pdf")

Marvinbryantia.ps <- prune_taxa("a1fe1f45720a7ea7649cbaff45fa34e3", DSSx.rel3) # 6 taxa in this genera, doing the sepcific asv
Marvinbryantia <- psmelt(Marvinbryantia.ps) 
RA.plot(Marvinbryantia, "Clostridia, Marvinbryantia", "Marvinbryantia.pdf") #doesn't look like much, let's find the specific otu



DSSxtg.RA <- transform(DSSxtg, "compositional")


tax_table(WC)[51]
attr(WC, "original_names")[51]
"b1d42839235884b6341cc6faed8133c8"

prune_taxa(attr(WC, "original_names")[51], DSSxtg.RA) %>% psmelt() %>% RA.plot(., "Eubacterium", "Eubacterium.pdf")

subset_taxa(DSSxtg.RA, Family == "Peptococcaceae") %>%  tax_glom(., "Class") %>% psmelt() %>% RA.plot(., "Clostridia, Peptococcaceae", "Peptococcaceae.pdf")

subset_taxa(DSSxtg.RA, Class== "Alphaproteobacteria") %>%  tax_glom(., "Class") %>% psmelt() %>% RA.plot(., "Alphaproteobacteria", "Alphaproteobacteria.pdf")

subset_taxa(DSSxtg.RA, Family== "Erysipelotrichaceae") %>% tax_glom(., "Family") %>% psmelt() %>% RA.plot(., "Firmicutes, Erysipelotrichaceae ", "Erysipelotrichaceae .pdf")

subset_taxa(DSSxtg.RA, Genus== "Bifidobacterium") %>% psmelt() %>% RA.plot(., "Bifidobacterium", "Bifidobacterium2.pdf")

subset_taxa(DSSxtg.RA, Family== "Family XIII") %>% psmelt() %>% RA.plot(., "Clostridiales, Family XIII ", "Family XIII.pdf")

subset_taxa(DSSxtg.RA, Genus== "Pseudomonas") %>% psmelt() %>% RA.plot(., "Pseudomonas", "Pseudomonas.pdf")

subset_taxa(DSSxtg.RA, Genus== "Lactobacillus") %>% psmelt() %>% RA.plot(., "Lactobacillus", "Lactobacillus.pdf")

subset_taxa(DSSxtg.RA, Family== "Ruminococcaceae") %>% tax_glom(., "Family") %>% psmelt() %>% RA.plot(., "Clostridiales, Ruminococcaceae", "Ruminococcaceae.pdf")
```

#Figure 6b RA plots Colour-coding by signficiance 
```{r}
Am <- subset_taxa(DSSxtg.RA, Genus == "Akkermansia") %>% psmelt()

Am$Physiology <- factor(Am$Physiology, levels = c("PELLET", "BCpos", "HNA", "LNA", "PI")  )

ggplot(Am, aes(x = Day, y = Abundance)) + geom_point(alpha = 0.5, aes(color = Physiology)) + geom_smooth(method = "loess", aes(color = Physiology, linetype = Physiology)) + theme_bw() + xlim(c(-3,10)) + ylab("Relative Abundance") + ggtitle("Akkermansia muciniphila") + scale_linetype_manual(values=c("solid", "solid","solid","solid","dashed")) + theme(text = element_text(size = 16))  

ggsave("Amuciniphila.pdf", width = 6.8, height = 5.2)

#Erisy
Ery <- subset_taxa(DSSxtg.RA, Family== "Erysipelotrichaceae") %>% tax_glom(., "Family") %>% psmelt()
Ery$Physiology <- factor(Ery$Physiology, levels = c("PELLET", "BCpos", "HNA", "LNA", "PI")  )
ggplot(Ery, aes(x = Day, y = Abundance)) + geom_point(alpha = 0.5, aes(color = Physiology)) + geom_smooth(method = "loess", aes(color = Physiology, linetype = Physiology)) + theme_bw() + xlim(c(-3,10)) + ylab("Relative Abundance") + ggtitle("Erysipelotrichaceae") + scale_linetype_manual(values=c("solid", "dashed","dashed","dashed","dashed")) + theme(text = element_text(size = 16))  
ggsave("Erysipelotrichaceae.pdf", width = 6.8, height = 5.2)

#bifido
bifido <- subset_taxa(DSSxtg.RA, Genus== "Bifidobacterium") %>% psmelt()
bifido$Physiology <- factor(bifido$Physiology, levels = c("PELLET", "BCpos", "HNA", "LNA", "PI")  )
ggplot(bifido, aes(x = Day, y = Abundance)) + geom_point(alpha = 0.5, aes(color = Physiology)) + geom_smooth(method = "loess", aes(color = Physiology, linetype = Physiology)) + theme_bw() + xlim(c(-3,10)) + ylim(c(0,0.3)) + ylab("Relative Abundance") + ggtitle("Bifidobacteria") + scale_linetype_manual(values=c("dashed", "solid","dashed","dashed","dashed")) + theme(text = element_text(size = 16))  
ggsave("Bifidobacteria.pdf", width = 6.8, height = 5.2)

#FamXIII
FamXIII <- subset_taxa(DSSxtg.RA, Family== "Family XIII") %>% tax_glom(., "Family") %>% psmelt()
FamXIII$Physiology <- factor(FamXIII$Physiology, levels = c("PELLET", "BCpos", "HNA", "LNA", "PI")  )
ggplot(FamXIII, aes(x = Day, y = Abundance)) + geom_point(alpha = 0.5, aes(color = Physiology)) + geom_smooth(method = "loess", aes(color = Physiology, linetype = Physiology)) + theme_bw() + xlim(c(-3,10)) + ylim(c(0,0.03)) + ylab("Relative Abundance") + ggtitle("Clostridiales Family XIII") + scale_linetype_manual(values=c("dashed", "solid","dashed","solid","dashed")) + theme(text = element_text(size = 16))  
ggsave("Family XIII.pdf", width = 6.8, height = 5.2)
```

#Firmicutes
```{r}
#Since Firmicutes are much more diverse, they are also changing in abdunace in many different ways. let's look at this closer

Firmicutes <- subset_taxa(DSSx, Phylum == "Firmicutes")
                          
FHNA <- subset_samples(Firmicutes, Physiology == "HNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FLNA <- subset_samples(Firmicutes, Physiology == "LNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FPI <- subset_samples(Firmicutes, Physiology == "PI") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FBCpos <- subset_samples(Firmicutes, Physiology == "BCpos") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FWC <- subset_samples(Firmicutes, Physiology == "PELLET") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()

cc(FHNA, "FHNAcc.pdf")
cc(FLNA, "FLNAcc.pdf")
cc(FPI, "FPIcc.pdf")
cc(FBCpos, "FBCposcc.pdf")
cc(FWC, "FWCcc.pdf")  

Firm.Family <- tax_glom(Firmicutes, "Family")
FFHNA <- subset_samples(Firm.Family, Physiology == "HNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FFLNA <- subset_samples(Firm.Family, Physiology == "LNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FFPI <- subset_samples(Firm.Family, Physiology == "PI") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FFBCpos <- subset_samples(Firm.Family, Physiology == "BCpos") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FFWC <- subset_samples(Firm.Family, Physiology == "PELLET") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()

cc(FFHNA, "FFHNAcc.pdf")
cc(FFLNA, "FFLNAcc.pdf")
cc(FFPI, "FFPIcc.pdf")
cc(FFBCpos, "FFBCposcc.pdf")
cc(FFWC, "FFWCcc.pdf")  
```

#corncob phylum level
```{r}
Phylum <- tax_glom(DSSx, "Phylum")

PHNA <- subset_samples(Phylum, Physiology == "HNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
PLNA <- subset_samples(Phylum, Physiology == "LNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
PPI <- subset_samples(Phylum, Physiology == "PI") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
PBCpos <- subset_samples(Phylum, Physiology == "BCpos") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
PWC <- subset_samples(Phylum, Physiology == "PELLET") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()

cc(PHNA, "PHNAcc.pdf")
cc(PLNA, "PLNAcc.pdf")
cc(PPI, "FPIcc.pdf")
cc(PBCpos, "FBCposcc.pdf")
cc(PWC, "FWCcc.pdf")         

PhylumDS <- subset_samples(Phylum, DiseaseState == "baseline" | DiseaseState == "pre-symptoms") %>% clean_taxa_names()
daWC <- differentialTest(
  formula = ~ DiseaseState * Physiology ,
  phi.formula = ~ DiseaseState , 
  formula_null = ~ 1, 
  phi.formula_null = ~ DiseaseState , 
  test = "Wald", boot = FALSE, 
  data = PhylumDS,
  fdr_cutoff = 0.05)

plot(daWC, col = "Taxa")
heatmap(daWC)
daWC$significant_models
corncob::plot.bbdml(daWC)
corncob::plot.diff

plot.differentialTest(daWC)

#How are firmciutes behaving overall?
corncob_da <- bbdml(formula = OTU4 ~ DiseaseState,
phi.formula = ~ DiseaseState,
data = PHNA)
plot(corncob_da, color = "DiseaseState")
summary(corncob_da)

#family level
Family <- tax_glom(DSSx, "Family")

FamHNA <- subset_samples(Family, Physiology == "HNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FamLNA <- subset_samples(Family, Physiology == "LNA") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FamPI <- subset_samples(Family, Physiology == "PI") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FamBCpos <- subset_samples(Family, Physiology == "BCpos") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()
FamWC <- subset_samples(Family, Physiology == "PELLET") %>% subset_samples(., DiseaseState == "baseline" | DiseaseState == "pre-symptoms")  %>% clean_taxa_names()

cc(FamHNA, "FamHNAcc.pdf")
cc(FamLNA, "FamLNAcc.pdf")
cc(FamPI, "FamPIcc.pdf")
cc(FamBCpos, "FamBCposcc.pdf")
cc(FamWC, "FamWCcc.pdf") 
```

